---
title: "Used Car Resale Values"
author: "Anthony Parrillo"
date: "5/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Purpose 

-Which brands (makes) of cars hold their value the best? 

-What factors significantly alter resale values? 

# Motivation

-Nearly everyone buys a used car or sells a used car at some point, if not many times in one's lifetime. 

## Carfax

Carfax.com is a website that is well known for checking the history and status of used cars to help used car buyers from being sold a car that has problems that the buyer is unaware of.

The data was scraped from used cars with filters of:

- Sold within a *50 mile radius* of New York city

- Under *$15,000*
  
### Extraction

I used selenium to automatically select drop down menues and fill in values corresponding to the filters mentioned above. The end result was to extract URLs containing the deatiled information of each car. 

Once all the URLS were collected, I extracted the detailed information for each car using scrapy.

###Other factors

The prices are based on dealer selling prices.

In total, there are 6747 used car rows included in this analysis. 

```{r}
library(ggplot2)
library(dplyr)
library(RColorBrewer)

cardata = read.csv("carfax/finaldata.csv")
head(cardata)
#remove the commas
cardata$price = as.numeric(gsub(",", "", cardata$price))
cardata$mileage = as.numeric(gsub(",", "", cardata$mileage))
#Fix Land Rover Entries
cardata$make = gsub("Land", "Land Rover", cardata$make)
#Rename column to yr
names(cardata)[17] = 'yr'
#Set scale to not use sci notation
options(scipen=6)

#preparing values for the t.tests: 
withaccident = cardata$price[cardata$accident != "No accidents reported to CARFAX"]
withoutaccident = cardata$price[cardata$accident == "No accidents reported to CARFAX"]
withautomatic = cardata$price[cardata$transmission == "Automatic"]
withmanual = cardata$price[cardata$transmission == "Manual"]

#preparing grouping for the plots
yrgroup = cardata %>% group_by(yr) %>% filter(yr >= 2000) %>% summarize(tot = n(), meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))
mileagegroup = cardata %>% summarize(meanval = mean(mileage), medianval = median(mileage), std = sd(mileage))
acc = cardata %>% group_by(accident) %>% summarize(medianval = median(price), std = sd(price))
mkgroup = cardata %>% group_by(make) %>% summarize(medianval = median(price), std = sd(price)) %>% arrange(desc(medianval))
trnsmsngroup = cardata %>% group_by(transmission) %>% summarize(meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))
fuelgroup = cardata %>% group_by(fuel) %>% summarize(meanval = median(price), std = sd(price)) %>% arrange(desc(meanval))
bdygroup = cardata %>% group_by(body_style) %>% summarize(meanval = median(price), std = sd(price)) %>% arrange(desc(meanval))
modeldata = cardata %>% group_by(model) %>% filter(n()>=50) %>% summarize(tot = n(), meanval = median(price), std = sd(price)) %>% arrange(desc(meanval))

acc[is.na(acc)] = 0
#Pull out the no accidents line of data
noacc = (acc[16, 2:3])
acc = acc[c(1:15, 17:20), ]
accdata = acc %>% summarise(mn = median(medianval), newstd = sd(std))
#noacc gives side-by-side comparison of 
noacc[2, ] = c(accdata$mn, accdata$newstd)
noacc = noacc %>% mutate(accreport = c('No Accidents Reported', 'Accidents Reported'))
noacc = noacc %>% select(accreport, medianval, std)

#boxplot of the data by year
cardata %>% ggplot(., aes(x = yr, y = price)) + geom_jitter(color = "dark green") + xlab('Year') + ylab('Price') + ggtitle("Used Car Resale Data: Price vs. Year") + scale_colour_hue()
cardata %>% ggplot(., aes(x = reorder(yr, yr), y = price)) + geom_boxplot() + ggtitle("Used Car Resale Data: Price vs. Year") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Year') + ylab('Selling Price')
yrgroup %>% ggplot(., aes(x = reorder(yr, yr), y = meanval)) + geom_bar(stat = 'identity', fill = "orange")+ geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Year") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Year') + ylab('Avg. Selling Price')
yrgroup %>% ggplot(., aes(x = reorder(yr, yr), y = tot)) + geom_bar(stat = 'identity', fill = "purple") + ggtitle("Used Car Resale Data: Number of Listings by Year") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Year') + ylab('Number of Listings')

#Number of used cars vs. Mileage
cardata %>% ggplot(., aes(x = mileage)) + geom_histogram(bins = 35, fill = "blue") + ggtitle('Used Cars for Sale by Mileage (6747 total)') + xlab('Mileage') + ylab("Number For Sale") + geom_density(alpha = .2, fill = 'white')

plot(density(cardata$mileage), xlab = "mileage", main = "Density Plot of Used Cars for Sale by Mileage (6747 total)", col = "dark green")
abline(v = mean(cardata$mileage), lwd = 2, lty = 2, col = "purple")
abline(v = median(cardata$mileage), lwd = 2, lty = 2, col = "magenta")
legend("topright", c("Mean", "Median"), lwd = 2, lty = 2, col = c("purple", "magenta"))

mileagegroup
    
#density hex plot by mileage
cardata %>% ggplot(., aes(x = mileage, y = price)) + geom_hex() + ggtitle("Used Car Resale Data: Price vs. Mileage") + xlab('Mileage') + ylab('Selling Price')
cardata %>% ggplot(., aes(x = mileage, y = price)) + geom_density2d() + ggtitle("Used Car Resale Data: Price vs. Mileage") + xlab('Mileage') + ylab('Selling Price')

#mileage, year vs. price
cardata %>% ggplot(., aes(x = yr, y = mileage)) + geom_hex() + xlab('Year') + ylab('Mileage') + ggtitle("Used Car Resale Data: Mileage vs. Year")

#By Make
mkgroup %>% ggplot(., aes(x = reorder(make, desc(medianval)), y = medianval)) + geom_violin() + geom_errorbar(aes(ymin=medianval-std, ymax=medianval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Make") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Make') + ylab('Avg. Selling Price')
mkgroup = as.data.frame(mkgroup)
mkgroup

#By Body Style 
bdygroup %>% ggplot(., aes(x = reorder(body_style, desc(meanval)), y = meanval)) + geom_bar(stat = "identity", fill = "magenta") + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Body Style") + xlab('Body Style') + ylab('Avg. Selling Price')

#By Fuel Type
fuelgroup %>% ggplot(., aes(x = reorder(fuel, desc(meanval)), y = meanval)) + geom_bar(stat = "identity", fill = "green") + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Energy Source") + xlab('Energy Source') + ylab('Avg. Selling Price')

#By Transmission
trnsmsngroup %>% ggplot(., aes(x = transmission, y = meanval)) + geom_bar(stat = "identity", fill = "gray") + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Transmission Type") + xlab('Transmission') + ylab('Avg. Selling Price')

plot(density(withautomatic), xlab = "price", main = "Resale Value of Automatic vs. Manual Transmission", col = "orange")
lines(density(withmanual), col = "dark green")
legend("topleft", c("Automatic", "Manual"), lwd = 2, col = c("orange", "dark green"))
abline(v = mean(withautomatic), lwd = 2, lty = 2, col = "orange")
abline(v = mean(withmanual), lwd = 2, lty = 2, col = "dark green")

t.test(withautomatic, withmanual, alternative = "two.sided")

#By Accident
numacc = cardata %>% group_by(accident) %>% summarize(tot = n(), medianval = median(price), std = sd(price))
numacc %>% filter(tot > 7) %>% ggplot(., aes(x = reorder(accident, desc(medianval)), y = medianval)) + geom_bar(stat = "identity", fill = "brown") + geom_errorbar(aes(ymin=medianval-std, ymax=medianval+std), width=.1) + geom_point() + ggtitle("Comparing Used Cars by Title Status") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Title Status') + ylab('Selling Price')
noacc %>% ggplot(., aes(x = accreport, y = medianval)) + geom_bar(stat = "identity", fill = "gold") + geom_errorbar(aes(ymin=medianval-std, ymax=medianval+std), width=.1) + geom_point() + ggtitle("Side-by-Side Used Car Accident vs. Non-Accident Comparison") + xlab('') + ylab('Selling Price')

plot(density(withoutaccident), xlab = "price", main = "Accident vs. No Accident Data", col = "red")
lines(density(withaccident), col = "blue")
legend("topleft", c("No Accidents Reported", "Accidents Reported"), lwd = 1, col = c("red", "blue"))
abline(v = mean(withoutaccident), lwd = 2, lty = 2, col = "red")
abline(v = mean(withaccident), lwd = 2, lty = 2, col = "blue")

t.test(withaccident, withoutaccident, alternative = "two.sided")

#By Model
modeldata %>% ggplot(., aes(x = reorder(model, desc(meanval)), y = meanval)) + geom_bar(stat = 'identity', fill = 'cyan')+ geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Model (50 or More Data Points)") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Model') + ylab('Avg. Selling Price')

```

## Conclusions:

1. Sedans resell better than other body types of used cars. 
2. An automatic transmission resells at $1300 more than a manual transmission. 
3. A used car without accidents reported resells at $1350 more than a car with accidents reported. 

Interesting fact: Gold color cars resell the worst of all exterior colors. 

These results can be applied not only for those who are interested in purchasing a used car, but also for:

- Those considering purchasing a new car.
- Whether or not to buy or lease a new car.
- Whether to sell a currently owned used car or keep it. 