---
title: "Used Car Resale Values"
author: "Anthony Parrillo"
date: "5/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Carfax.com

Carfax.com is a website that is most well known for checking the history and status of used cars to help used car buyers from being sold a car that has problems that the buyer is unaware of.

The data was scraped from used cars with filters of being sold within a 50 mile radius of New York city, and under $15,000. The car URLs were extracted using selenium, and then the data for each specific car was taken using a scrapy spider. 

There are a total of 6747 used car rows included in this analysis. 

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(RColorBrewer)

cardata = read.csv("carfax/finaldata.csv")
#remove the commas
cardata$price = as.numeric(gsub(",", "", cardata$price))
cardata$mileage = as.numeric(gsub(",", "", cardata$mileage))
#Fix Land Rover Entries
cardata$make = gsub("Land", "Land Rover", cardata$make)
#Rename column to yr
names(cardata)[17] = 'yr'
#Set scale to not use sci notation
options(scipen=6)

#preparing grouping for the plots
yrgroup = cardata %>% group_by(yr) %>% summarize(meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))
mileagegroup = cardata %>% group_by(mileage) %>% summarize(meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))
acc = cardata %>% group_by(accident) %>% summarize(meanval = mean(price), std = sd(price))
mkgroup = cardata %>% group_by(make) %>% summarize(meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))
trnsmsngroup = cardata %>% group_by(transmission) %>% summarize(meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))
fuelgroup = cardata %>% group_by(fuel) %>% summarize(meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))
bdygroup = cardata %>% group_by(body_style) %>% summarize(meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))
modeldata = cardata %>% group_by(model) %>% filter(n()>=50) %>% summarize(number = n(), meanval = mean(price), std = sd(price)) %>% arrange(desc(meanval))

acc[is.na(acc)] = 0
#Pull out the no accidents line of data
noacc = (acc[16, 2:3])
acc = acc[c(1:15, 17:20), ]
accdata = acc %>% summarise(mn = mean(meanval), newstd = sd(std))
#noacc gives side-by-side comparison of 
noacc[2, ] = c(accdata$mn, accdata$newstd)
noacc = noacc %>% mutate(accreport = c('No Accidents Reported', 'Accidents Reported'))
noacc = noacc %>% select(accreport, meanval, std)

#boxplot of the data by year
cardata %>% ggplot(., aes(x = yr, y = price)) + geom_jitter() + xlab('Year') + ylab('Price') + ggtitle("Used Car Resale Data: Price vs. Year")
cardata %>% ggplot(., aes(x = reorder(yr, yr), y = price)) + geom_boxplot() + ggtitle("Used Car Resale Data: Price vs. Year") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Year') + ylab('Selling Price')
yrgroup %>% ggplot(., aes(x = reorder(yr, yr), y = meanval)) + geom_bar(stat = 'identity')+ geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Year") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Year') + ylab('Avg. Selling Price')

#Number of used cars vs. Mileage
cardata %>% ggplot(., aes(x = mileage)) + geom_histogram(bins = 40) + ggtitle('Used Cars for Sale by Mileage (6747 total)') + xlab('Mileage') + ylab("Number For Sale") + geom_density(alpha = .2, fill = 'white')
cardata %>% ggplot(., aes(x = mileage)) + geom_density() + ggtitle('Used Cars for Sale by Mileage (6747 total)') + xlab('Mileage') + ylab("Density")
cardata %>% ggplot(., aes(x = yr, y = mileage)) + geom_hex() + xlab('Year') + ylab('Mileage') + ggtitle("Used Car Resale Data: Mileage vs. Year")

#density hex plot by mileage
cardata %>% ggplot(., aes(x = mileage, y = price)) + geom_hex() + ggtitle("Used Car Resale Data: Price vs. Mileage") + xlab('Mileage') + ylab('Selling Price')

#By Make
mkgroup %>% ggplot(., aes(x = reorder(make, desc(meanval)), y = meanval)) + geom_violin() + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Make") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Make') + ylab('Avg. Selling Price')
mkgroup 

#By Body Style 
bdygroup %>% ggplot(., aes(x = reorder(body_style, desc(meanval)), y = meanval)) + geom_bar(stat = "identity") + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Body Style") + xlab('Body Style') + ylab('Avg. Selling Price')

#By Fuel Type
fuelgroup %>% ggplot(., aes(x = reorder(fuel, desc(meanval)), y = meanval)) + geom_bar(stat = "identity") + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Fuel Type") + xlab('Fuel Type') + ylab('Avg. Selling Price')

#By Transmission
trnsmsngroup %>% ggplot(., aes(x = transmission, y = meanval)) + geom_bar(stat = "identity") + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Transmission Type") + xlab('Transmission') + ylab('Avg. Selling Price')
trnsmsngroup

#By Accident
acc %>% ggplot(., aes(x = reorder(accident, desc(meanval)), y = meanval)) + geom_bar(stat = "identity") + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Comparing Used Cars by Title Status") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Title Status') + ylab('Selling Price')
noacc %>% ggplot(., aes(x = accreport, y = meanval)) + geom_bar(stat = "identity") + geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Side-by-Side Used Car Accident vs. Non-Accident Comparison") + xlab('') + ylab('Selling Price')
noacc

#By Model
modeldata %>% ggplot(., aes(x = reorder(model, desc(meanval)), y = meanval)) + geom_bar(stat = 'identity')+ geom_errorbar(aes(ymin=meanval-std, ymax=meanval+std), width=.1) + geom_point() + ggtitle("Used Car Resale Data: Price vs. Model (50 or More Data Points)") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab('Model') + ylab('Avg. Selling Price')
```